Performance Evaluation of Big Data Processing Strategies for Neuroimaging

### 2018-12-13 paper.tex
Complete read and editing

\usepackage{graphicx}
\usepackage{color}
\usepackage{colortbl} % for \rowcolor
\usepackage{caption}
\usepackage{subcaption}
\usepackage[bookmarks=false]{hyperref}
\usepackage{algpseudocode}
\usepackage{algorithm}
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}

@@

@@

Big Data processing engines reduce data-related overheads, through 
use of concepts such as data locality, in-memory computing, and lazy evaluation.
Data locality-aware scheduling, popularized by the MapReduce~\cite{dean2008mapreduce} 
framework, runs tasks to the nodes holding the data rather than 
performing costly data transfers across the network. In-memory 
computing, adopted by MapReduce's successor Apache 
Spark~\cite{zaharia2016apache}, caches data in memory whenever 
possible, reducing costly disk I/Os between pipeline steps. 
Lazy evaluation, also used in Spark, means that results are computed only 
when necessary, enabling further performance optimizations. Frameworks 
such as
MapReduce and Spark have become mainstream tools for data analytics, 
although many others, such as 
Dask~\cite{rocklin2015dask}, are emerging. 
Meanwhile, several scientific 
domains including bioinformatics, physics or astronomy, have entered 
the Big Data era due to increasing data volumes and variety. 
However, the adoption of Big Data engines for scientific data analysis 
remains limited, perhaps due to the widespread availability of 
scientific processing engines such as Pegasus~\cite{deelman2005pegasus} or
Taverna~\cite{oinn2004taverna}, and the adaptations required in Big 
Data processing engines for scientific computing.

Scientific applications differ from the typical Big Data use 
cases, which might explain the remaining gap between Big Data and 
scientific engines. \todo{This sentence is extremely long} While Big Data applications mostly target text 
processing (e.g. Web search, frequent pattern mining, recommender 
systems~\cite{leskovec2014mining}), implemented in consistent software 
libraries and running on clouds or dedicated commodity clusters with 
locality-aware file systems such as the Hadoop Distributed File System 
(HDFS~\cite{shvachko2010hadoop}), scientific applications often involve 
binary data such as images and signals, processed by command-line tools 
using a mix of programming languages (C, Fortran, Python, shell 
scripts), and are deployed on large, shared clusters where 
data is fetched from data nodes to compute nodes using centralized shared file 
systems such as Lustre~\cite{schwan2003lustre}. Such differences in 
applications and infrastructures have important consequences. To 
mention only one, in-memory computing requires instrumentation to be 
applied to command-line tools. Moreover, performance is only one of the important
aspects to scientific computing. Other important aspects include reproducibility 
and provenance tracking. While these are minimally addressed by Big Data
processing engines, it is currently insufficient for scientific computing. 
Incorporation of provenance techniques into frameworks such are Spark are
currently being investigated~\cite{samba}.

Technological advances of the past decade, in particular page caching 
in the Linux kernel~\cite{love2010linux}, in-memory file systems 
(tmpfs) and memory-mapped files\todo{check that} might also 
explain the lack of adoption of Big Data engines for scientific 
applications.
In-memory computing would then be a feature provided by 
the operating system rather than by the engine itself. The frontier 
between these two components is blurred and needs to be clarified.

@@

Answers to these questions have important implications. 
In~\cite{mehta2017comparative}, a comparative study of Dask, Spark, 
TensorFlow, MyriaDB, and SciDB on neuroinformatics use-cases is 
presented. It concludes that these systems need to be extended to 
better address scientific code integration, data partitioning, data 
formats and system tuning. We argue 
that such efforts should only be conducted if substantial performance 
improvements are expected from in-memory computing, lazy 
evaluation\todo{are we talking about lazy evaluation at all?} or data 
locality. On the other hand, neuroimaging data processing engines are 
still being developed, and it is legitimate to wonder whether these 
projects should just be migrated to Spark, Dask, or other Big Data 
engines.

Our study focuses on performance. We intentionally do not intend to 
compare Big Data and scientific data processing engines on the grounds 
of workflow language expressivity, fault-tolerance, provenance capture and 
representation, portability or reproducibility, which are otherwise 
critical concerns. Besides, our study of performance focuses on the 
impact of data writes and transfers. It purposely leaves out task 
scheduling to computing resources, to focus on the understanding of 
data writes and movement. Task scheduling will be part of our 
discussion though.

Infrastructure-wise, we focus on the case of High-Performance Computing 
(HPC) clusters that are typically available through University 
facilities or national computing infrastructures such as XSEDE, Compute 
Canada or DEISA \todo{add links}, as this is the type of platform which 
is typically available to neuroscientists. We assume that HPC systems 
are multi-tenant, that compute nodes are accessible through a batch 
scheduler, and that a file system shared among the compute nodes is 
available, for instance Lustre. We intentionally leave distributed, 
HDFS-like file system out, as initiatives to deploy them in HPC 
centers, for instance Hadoop on-demand~\cite{hadoop-on-demand} have not 
become mainstream yet.

@@

Our methods, including performance models, processing engines, \todo{Make this paragraph more informative}
applications and infrastructure used are described in 
Section~\ref{sec:methods}. Section~\ref{sec:results} presents our 
results which we discuss in Section~\ref{sec:discussion}. 
Section~\ref{sec:conclusion} concludes on the answer to our two research questions.

@@

\todo{Maybe add an intro to this section, to outline it}

@@

An RDD is an immutable parallel data structure that achieve fault-tolerance 
through the concept of lineage~\cite{zaharia2010spark}. Rather than permitting
fine-grained transformations, only coarse-grained transformations (applying to
many elements in the RDD), can be applied, thereby making it simple to keep a 
log of how the data was modified. This log, known as the lineage, is used
to reproduce any lost data.

Two types of operations can be performed on RDDs:
transformations and actions. Applying a transformation to an RDD produces a new,
child, RDD through a narrow or wide dependency. A narrow dependency signifies 
that the child is only dependent on a single parent partition, whereas a child 
RDD is dependent on all parent partitions in a wide dependency. Examples of 
Spark transformations include map, filter and join. To materialize an RDD, an
action must be performed, such as a reduce or a collect.

@@

Executing Spark applications on HPC systems with Spark-unaware schedulers may
be inefficient. The amount of resources requested by Spark may impede Spark-cluster
scheduling time. Using pilot-scheduling strategies to add nodes to the Spark cluster
as they are allocated by the underlying schedulers may speedup allocation and 
overall processing time. This, however, is not studied in the current paper.

As Spark is frequently used by the scientific community, we designed our 
experiments using the PySpark API. This is at a cost to performance as the PySpark
code must undergo Python to Java serialization.

@@

@@

Data storage location is critical to the performance of Big Data 
applications on HPC clusters.
Data may reside in the engine memory, on 
a file system whose content reside in virtual memory (for instance 
tmpfs), on disks local to the processing node, or on a shared 
file system. Table~\ref{table:features} summarizes the Big Data 
strategies that can be used depending on the data location. In 
addition, lazy evaluation is available in Spark regardless of data 
location. The remainder of this Section explains this Table and 
provides related performance models.
\begin{table}
\centering
\begin{tabular}{c|cc}
   \rowcolor{headcolor}
    Data Location                 & In-Memory     & Data Locality        \\
    \rowcolor{headcolor}
                                  & Computing     &                     \\
                                  \hline          
In-memory                         &   Y           & Y                       \\
tmpfs                             &   Y           & Y                   \\
Local Disk                        & Page Caching  & Y                   \\
Shared File System                & Page Caching  & N                
% data locality is somehow present in shared fs due to page caching again.
\end{tabular}
\caption{Big Data strategies on a shared HPC cluster.}
\label{table:features}
\end{table}

@@

 The main difference between storing the data in the engine memory, as in Spark, 
 and simply writing to an in-memory file system, such as tmpfs, is 
 related to what happens when the processed data fills up the available 
 memory. In case engine memory is used, the engine has to cleanup 
 unused data with some strategy to avoid crashes. In case an in-memory 
 file system is used, the kernel starts swapping memory to disk, and 
 the performance becomes that of local disk writes. In our experiments, 
 we will explore the region  when the data consumed by the 
 application approaches the threshold of available memory.

\subsubsection{Local Disk (Page Caching)} % 0.5 page

@@

Storing data on local disks obviously enables data locality as tasks 
executed on the node where their input files reside do not require to 
transfer input or output data. However, in absence of a more specific 
file system such as the Hadoop Distributed File System (HDFS) to handle 
file replication across computing nodes, data locality comes at the price
of stringent scheduling restrictions, as tasks can only be scheduled to the
single node that contains their input data.
% this brings data locality but not at the level of HDFS because replication isn't there and 
% there is only 1 single option to schedule a task while preserving data locality.

The performance of local disk accesses is strongly dependent on the 
page caching mechanism provided by the Linux kernel, described in    
details in~\cite{love2010linux}. To summarize, data read from disk 
remains cached in memory until evicted by an LRU (Least Recently Used) 
strategy. When a process invokes the \texttt{read()} system call, the 
kernel will return the data directly from memory if the requested data 
lies in the page cache, realizing a cache \emph{hit}. Cache hits drastically speed-up data 
reads, by masking the disk latency and limited bandwidth behind a 
memory buffer. In effect, page caching provides in-memory computing 
transparently to the processing engine. However, page cache eviction 
strategies currently cannot be controlled by the application 
\todo{check that}, which prevents processing engines to anticipate 
reads by preloading the cache. Scheduling strategies might be designed 
to maximize cache hits though. For instance, lazy 
evaluation could result in more cache hits by scheduling data-dependent 
tasks on the same node.

@@

We introduce the following basic model to describe the filling and 
flushing of the page cache by an application:
$$
d(t) = \left( \frac{D}{C} - \frac{\delta}{\gamma} \right)t + d_0,
$$
where:
\begin{itemize}
\item $d(t)$ is the amount of data in the page cache at time t
\item $D$ is the total amount of data written by the application
\item $C$ is the total CPU time of the application
\item $\delta$ is the disk bandwidth
\item $\gamma$ is the max number of concurrent processes
\item $d_0$ is the amount of data in the page cache at time $t_0$
\end{itemize}

@@

In general, an application should ensure that $\dot d$ remains negative 
or null, leading to the following inequality:
\begin{equation}
\frac{D}{C} \leq \frac{\delta}{\gamma} \label{eq:page-cache-inequality}
\end{equation}
This defines a data-compute ratio beyond which the page cache becomes 
asymptotically useless. We will use this ratio to define our benchmarks. It 
should be noted that leveraging the page cache, and therefore ensuring 
that Equation~\ref{eq:page-cache-inequality} holds, has important 
performance implications: with page caching, the application will 
experience negligible write times, while without page caching 
the write throughput will be that of the disk.

@@

\todo{How about reads?}

@@

We model a shared file system using its global apparent bandwidth 
$\Delta$, shared by all concurrent processes in the system. We are 
aware that such a simplistic model does not describe at all the 
intricacies of systems such as Lustre. In particular, metadata 
management, RPC protocol optimizations and storage optimizations are 
all covered under the apparent bandwidth. We do, however, consider the 
effect of page caching in shared file systems too, since in Linux 
writes to network-mounted volumes benefit from this feature too.

As in the local disk model, we note that page caching will only be 
useful when the flush bandwidth is greater than the write throughput of 
the application, that is:
\begin{equation}
\frac{D}{C} \leq \frac{\Delta}{\Gamma}, \label{eq:page-cache-sharedfs}
\end{equation}
where $\Gamma$ is the max number of concurrent jobs in the cluster. 
Note that $\frac{\Delta}{\Gamma}$ will be much lower than 
$\frac{\delta}{\gamma}$ in Equation~\ref{eq:page-cache-inequality}.

@@

\subsection{Experiments} % 1.5 pages

\subsubsection{Infrastructure} % 1/4 a page (half a col)

All experiments were executed on the Dell Zenith cluster~\cite{dellHPC} running 
SLURM. For the 
Spark experiments, a Spark cluster was started atop our SLURM allocation. We
had access to 16 dedicated nodes. Each compute node ran on RHEL 7.4, was 
equipped with 40 Intel Xeon Gold 6148 2.4GHz CPUs and had
187GB of memory, 94GB of tmpfs and a Dell 120GB M.2 SATA SSD as local disk. 
A 698TB Lustre server is accessible on each node through a 100Gb/s Infiniband
network. The apparent bandwidth of the local disk was found to be 1480MB/s, 
whereas Lustre's bandwidth was found to be at 401.49MB/s.
% Dedicated Compute nodes with SLURM
% Lustre server

@@

\subsubsection{Datasets} % 1/4 of a page (half a col)

% BigBrain
% Some BIDS dataset for fmriprep
We used BigBrain~\cite{bigbrain}, a 75GB 40$\mu$m isotropic 
histological image of a 65-year-old male's brain. The BigBrain was selected due 
to being the only one of its kind, as there does not yet exist a 
higher-resolution image of a human brain. Moreover, there currently exists a 
lack of standardized tools for processing the BigBrain as a consequence 
of its size. To examine the effects processing BigBrain has on the Big Data 
strategies, we partitioned the full $3845\times3015\times3470$ voxel image into 30 ($5\times3\times2$)
chunks, 125 ($5\times5\times5$) chunks and 750 ($5\times15\times10$) chunks. Additionally, the full 
image was also split in half ($769\times603\times347$ voxels) and the half image was 
partitioned into 125 chunks.

Processing large images are only considered to be part of the BigData problem 
in neuroscience. The other problem is processing large MRI datasets, that is,
datasets consisting of many small brain images belonging to various different 
subjects. Although we have not explored explicitly the processing of large MRI
datasets, the 75GB BigBrain is within the size ballpark~\cite{VanHorn2014} of 
what a large MRI dataset would be today.

As it is possible that both small and large data may need to be processed using
the same pipeline \todo{explain pipelines in intro if not done already}, we 
wanted to determine the effects of the data management strategies on small 
data as well. For this, we selected a 12MB T1W image belonging to subject 1 of 
the ds000001 dataset version 6 ~\cite{openneuro}. In order to be able to split 
it into 125 chunks, it was necessary to zero-pad the image to the dimensions
$165\times200\times200$ voxels, which subsequently increased the total image size to 13MB.

\subsubsection{Applications} % 1 page
\begin{algorithm}\caption{Incrementation}\label{alg:incrementation}
    \begin{algorithmic}
    \Require $x$, a sleep delay in float
    \Require $fs$, filesystem to save image to (i.e. mem, tmpfs, local disk, lustre). 
    \ForEach{$i \in iterations$}
    \ForEach{$chunk \in image$}
        \State $chunk\gets chunk+1$
        \State sleep $x$
        \State save $chunk$ to $fs$
    \EndFor
    \EndFor
\end{algorithmic}
\end{algorithm}

\begin{figure*}
\begin{subfigure}{0.33\linewidth}
\centering
    \includegraphics[width=\textwidth]{results/figures/tmpfs-incrementation.pdf}
\caption{}
\label{fig:sub1}
\end{subfigure}%
    \begin{subfigure}{0.33\linewidth}
\centering
    \includegraphics[width=\textwidth]{results/figures/local-incrementation.pdf}
\caption{}
\label{fig:sub2}
    \end{subfigure}%
\begin{subfigure}{0.33\linewidth}
\centering
    \includegraphics[width=\textwidth]{results/figures/lustre-incrementation.pdf}
\caption{}
\label{fig:sub3}
\end{subfigure}
    \caption{Model evaluation on (a) tmpfs, (b) local and (c) Lustre. Grey 
             regions denote areas that violate model predictions.}
\label{fig:modeleval}
\end{figure*}

% Incrementation (avoid cache effects in binarization)
In order to investigate how the different strategies impact processing, we 
selected a simple incrementation pipeline (Algorithm~\ref{alg:incrementation}) 
that consisted exclusively of map 
stages. A series of map-only stages would enable us to evaluate the effects of
in-memory computing when data locality is preserved. Incrementation was 
selected 
over other applications, such as binarization, as it ensured that a new image
was created at each step (i.e. no caching effects within the executing 
application). Each partitioned chunk was incremented by 1, in parallel, by
the processing engines. As incrementing the images is a relatively quick 
process, we added a sleep delay to the incrementation stage in order to 
increase task duration. The incremented chunks would be either 
maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
Lustre (Spark and Nipype). Should more than a single iteration be requested, 
the increment chunks would be incremented again and saved to the same storage. 
This would repeat until the number of requested iterations had elapsed. In all 
conditions, the first input chunks would be reads from Lustre and the final 
output chunks would be saved to Lustre. We chose to perform our initial 
reads and final writes on Lustre as a multi-tenant clusters' compute node's 
local storage, if 
available, is generally temporary and only available for the duration of the 
allocation.

\subsubsection{Setup}

>>>>

\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{colortbl} % for \rowcolor
\usepackage{caption}
\usepackage{subcaption}
\usepackage[bookmarks=false]{hyperref}
\usepackage{algorithm}
\usepackage{algorithmicx} % Doc is at http://tug.ctan.org/macros/latex/contrib/algorithmicx/algorithmicx.pdf
\usepackage{algpseudocode}
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
\usepackage[bookmarks=false]{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}

@@

\algblock{Input}{EndInput}
\algnotext{EndInput}
\algblock{Output}{EndOutput}
\algnotext{EndOutput}
\newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}

@@

Big Data processing engines reduce data-related overheads, through 
use of concepts such as data locality, in-memory computing, and lazy evaluation.
Data locality-aware scheduling, popularized by the MapReduce~\cite{dean2008mapreduce} 
framework, runs tasks to the nodes holding the data rather than 
performing costly data transfers across the network. In-memory 
computing, adopted by MapReduce's successor Apache 
Spark~\cite{zaharia2016apache}, stores data in memory whenever 
possible, reducing costly disk I/Os between pipeline steps. 
Lazy evaluation, also used in Spark, means that results are computed only 
when necessary, enabling further performance optimizations. Frameworks 
such as
MapReduce and Spark have become mainstream tools for data analytics, 
although many others, such as 
Dask~\cite{rocklin2015dask}, are emerging. 
Meanwhile, several scientific 
domains including bioinformatics, physics or astronomy, have entered 
the Big Data era due to increasing data volumes and variety. 
However, the adoption of Big Data engines for scientific data analysis 
remains limited, perhaps due to the widespread availability of 
scientific processing engines such as Pegasus~\cite{deelman2005pegasus} or
Taverna~\cite{oinn2004taverna}, and the adaptations required in Big 
Data processing engines for scientific computing.

Scientific applications differ from the typical Big Data use 
cases, which might explain the remaining gap between Big Data and 
scientific engines. While Big Data applications mostly target text 
processing (e.g. Web search, frequent pattern mining, recommender 
systems~\cite{leskovec2014mining}) implemented in consistent software 
libraries, scientific applications often involve 
binary data such as images and signals, processed by command-line tools 
using a mix of programming languages (C, Fortran, Python, shell 
scripts). Infrastructure-wise, Big Data applications commonly run on 
clouds or dedicated commodity clusters with locality-aware file systems 
such as the Hadoop Distributed File System 
(HDFS~\cite{shvachko2010hadoop}), whereas scientific applications are 
usually deployed on large, shared clusters where data is moved between
data nodes and compute nodes using shared file systems such 
as Lustre~\cite{schwan2003lustre}. Such differences in applications and 
infrastructures have important consequences. To 
mention only one, in-memory computing requires instrumentation to be 
applied to command-line tools.

Technological advances of the past decade, in particular page caching 
in the Linux kernel~\cite{love2010linux}, in-memory file systems 
(tmpfs) and memory-mapped files might also 
explain the lack of adoption of Big Data engines for scientific 
applications.
In-memory computing would then be a feature provided by 
the operating system rather than by the engine itself. The frontier 
between these two components is blurred and needs to be clarified.

@@

Answers to these questions have important implications. 
In~\cite{mehta2017comparative}, a comparative study of Dask, Spark, 
TensorFlow, MyriaDB, and SciDB on neuroinformatics use-cases is 
presented. It concludes that these systems need to be extended to 
better address scientific code integration, data partitioning, data 
formats and system tuning. We argue 
that such efforts should only be conducted if substantial performance 
improvements are expected from in-memory computing, lazy 
evaluation or data 
locality. On the other hand, neuroimaging data processing engines are 
still being developed, and it is legitimate to wonder whether these 
projects should just be migrated to Spark, Dask, or other Big Data 
engines.

Our study focuses on performance. We intentionally do not intend to 
compare Big Data and scientific data processing engines on the grounds 
of workflow language expressivity, fault-tolerance, provenance capture 
and representation, portability or reproducibility, which are otherwise 
critical concerns, addressed for instance in~\cite{samba}. Besides, our 
study of performance focuses on the impact of data writes and 
transfers. It purposely leaves out task scheduling to computing 
resources, to focus on the understanding of 
data writes and movement. Task scheduling will be part of our 
discussion though.

Infrastructure-wise, we focus on the case of High-Performance Computing 
(HPC) clusters that are typically available through University 
facilities or national computing infrastructures such as 
\href{xsede.org}{XSEDE}, \href{http://computecanada.ca}{Compute Canada} 
or \href{http://www.prace-ri.eu}{PRACE}, as neuroscientists typically use such platforms.
 We assume that HPC systems are 
multi-tenant, that compute nodes are accessible through a batch 
scheduler, and that a file system shared among the compute nodes is 
available. We intentionally leave distributed, 
HDFS-like file system out, as initiatives to deploy them in HPC 
centers, for instance Hadoop on-demand~\cite{krishnan2011myhadoop}, have not 
become mainstream yet.

@@

Our methods, including performance models, processing engines, 
applications and infrastructure used are described in 
Section~\ref{sec:methods}. Section~\ref{sec:results} presents our 
results which we discuss in Section~\ref{sec:discussion} along with the 
two research questions mentioned previously. 
Section~\ref{sec:conclusion} concludes on the relevance of Big Data 
processing strategies for neuroimaging applications.

@@

The pipelines, benchmarks, performance data, and analysis scripts used 
to implement the methods described hereafter are all available at 
\url{https://github.com/big-data-lab-team/paper-in-mem-locality} for 
further inspection and reproducibility. Links to the pipeline engines 
and processed data are provided in the text.

@@

An RDD is an immutable parallel data structure that achieves fault-tolerance 
through the concept of lineage~\cite{zaharia2010spark}. Rather than permitting
fine-grained transformations, only coarse-grained transformations (applying to
many elements in the RDD), can be applied, thereby making it simple to keep a 
log of how the data was modified. This log, known as the lineage, is used
to reproduce any lost data.

Two types of operations can be performed on RDDs:
transformations and actions. Applying a transformation to an RDD produces a new,
child, RDD through a narrow or wide dependency. A narrow dependency signifies 
that the child is only dependent on a single parent partition, whereas a child 
RDD is dependent on all parent partitions in a wide dependency. Examples of 
Spark transformations include map, filter and join. To materialize an RDD, an
action must be performed, such as a reduce or a collect. \todo{Lazy evaluation should be mentioned here, or elsewhere in the paragraph.
It is currently not described.}

@@

Executing Spark applications on HPC systems with Spark-unaware schedulers may
be inefficient. The amount of resources requested by Spark may impede 
Spark-cluster scheduling time. Using pilot-scheduling strategies to add 
nodes to the Spark cluster as they are allocated by the underlying 
schedulers may speedup allocation and overall processing time 
\todo{Cite Shantenu's paper here}. This, however, is not studied in the 
current paper.

As Spark is frequently used by the scientific community, we designed 
our experiments using the PySpark API. This is at a cost to performance 
as the PySpark code must undergo Python to Java serialization. We used 
version 2.3.2 installed from \url{https://spark.apache.org}.

@@

We used Nipype version 1.1.4 installed through the pip package manager.

@@

Data storage location is critical to the performance of Big Data 
applications on HPC clusters.
Data may reside in the engine memory, on 
a file system whose content reside in virtual memory (for instance 
tmpfs), on disks local to the processing node, or on a shared 
file system. Table~\ref{table:features} summarizes the Big Data 
strategies that can be used depending on the data location. In 
addition, lazy evaluation is available in Spark regardless of data 
location. The remainder of this Section explains this Table and 
provides related performance models.
\begin{table}
\centering
\begin{tabular}{c|cc}
   \rowcolor{headcolor}
    Data Location                 & In-Memory     & Data Locality        \\
    \rowcolor{headcolor}
                                  & Computing     &                     \\
                                  \hline          
In-memory                         & \cellcolor{green!25} Yes           & \cellcolor{green!25}Yes                      \\
tmpfs                             & \cellcolor{green!25} Yes           & \cellcolor{green!25}Yes                  \\
Local Disk                        & \cellcolor{orange!25}Page Caching  & \cellcolor{green!25}Yes                  \\
Shared File System                & \cellcolor{orange!25}Page Caching  & \cellcolor{red!25}No                
% data locality is somehow present in shared fs due to page caching again.
\end{tabular}
\caption{Big Data strategies on a shared HPC cluster.}
\label{table:features}
\end{table}

@@

 The main difference between storing the data in the engine memory, as in Spark, 
 and simply writing to an in-memory file system, such as tmpfs, is 
 related to what happens when the processed data fills up the available 
 memory. In case engine memory is used, the engine must cleanup 
 unused data with some strategy to avoid crashes. In case an in-memory 
 file system is used, the kernel starts swapping memory to disk, and 
 the performance becomes that of local disk writes. In our experiments, 
 we will explore the region  when the data consumed by the 
 application approaches the threshold of available memory.

\subsubsection{Local Disk} % 0.5 page

@@

Storing data on local disks obviously enables data locality as tasks 
executed on the node where their input files reside do not require to 
transfer input or output data. However, in absence of a more specific 
file system such as HDFS to handle 
file replication across computing nodes, data locality comes at the price
of stringent scheduling restrictions, as tasks can only be scheduled to the
single node that contains their input data.
% this brings data locality but not at the level of HDFS because replication isn't there and 
% there is only 1 single option to schedule a task while preserving data locality.

The performance of local disk accesses is strongly dependent on the 
page caching mechanism provided by the Linux kernel, described in    
details in~\cite{love2010linux}. To summarize, data read from disk 
remains cached in memory until evicted by an LRU (Least Recently Used) 
strategy. When a process invokes the \texttt{read()} system call, the 
kernel will return the data directly from memory if the requested data 
lies in the page cache, realizing a cache \emph{hit}. Cache hits drastically speed-up data 
reads, by masking the disk latency and bandwidth behind a 
memory buffer. In effect, page caching provides in-memory computing 
transparently to the processing engine. However, page cache eviction 
strategies currently cannot be controlled by the application, which 
prevents processing engines to anticipate reads by preloading the 
cache. Scheduling strategies might be designed 
to maximize cache hits though. For instance, lazy 
evaluation could result in more cache hits by scheduling data-dependent 
tasks on the same node.

@@

We introduce the following basic model to describe the filling and 
flushing of the page cache by an application:
$$
d(t) = \left( \frac{D}{C} - \frac{\delta}{\gamma} \right)t + d_0,
$$
where:
\begin{itemize}
\item $d(t)$ is the amount of data in the page cache at time t
\item $D$ is the total amount of data written by the application
\item $C$ is the total CPU time of the application
\item $\delta$ is the disk bandwidth
\item $\gamma$ is the max number of concurrent processes on a node
\item $d_0$ is the amount of data in the page cache at time $t_0$
\end{itemize}

@@

In general, an application should ensure that $\dot d$ remains negative 
or null, leading to the following inequality:
\begin{equation}
\frac{D}{C} \leq \frac{\delta}{\gamma} \label{eq:page-cache-inequality}
\end{equation}
This defines a data-compute ratio beyond which the page cache becomes 
asymptotically useless. It should be noted that the transient phase 
during the page cache fills up might last a significant amount of time, 
in particular when $\dot d$ is positive but small. We intentionally do not model the 
transient phase as it requires detailed knowledge on the size of the 
page cache and the initial amount 
of data in it ($d_0$), which are difficult to estimate.

 We will use Equation~\ref{eq:page-cache-inequality} to 
define our benchmarks and interpret the results. It 
should be noted that leveraging the page cache, and therefore ensuring 
that Equation~\ref{eq:page-cache-inequality} holds, has important 
performance implications: with page caching, the write throughput will 
be that of memory, while without page caching it will be that of the 
disk.

@@

@@

We model a shared file system using its global apparent bandwidth 
$\Delta$, shared by all concurrent processes in the cluster. We are 
aware that such a simplistic model does not describe at all the 
intricacies of systems such as Lustre. In particular, metadata 
management, RPC protocol optimizations and storage optimizations are 
all covered under the apparent bandwidth. We do, however, consider the 
effect of page caching in shared file systems too, since in Linux 
writes to network-mounted volumes benefit from this feature too.

As in the local disk model, we note that page caching will only be asymptotically
useful when the flush bandwidth is greater than the write throughput of 
the application, that is:
\begin{equation}
\frac{D}{C} \leq \frac{\Delta}{\Gamma}, \label{eq:page-cache-sharedfs}
\end{equation}
where $\Gamma$ is the max number of concurrent processes in the cluster. 
Note that $\frac{\Delta}{\Gamma}$ will usually be much lower than 
$\frac{\delta}{\gamma}$.

@@

\subsection{Infrastructure} % 1/4 a page (half a col)

 All experiments were executed on  
 \href{https://www.dellemc.com/resources/en-us/asset/sales-documents/products/storage/h16221-hpc-lab-brochure.pdf}{Dell EMC's Zenith 
 cluster}, a Top-500 machine in the Dell EMC HPC Innovation Lab, running 
 SLURM. For the Spark experiments, a Spark cluster was started atop our 
 SLURM allocation. We had access to 16 dedicated nodes. Each compute 
 node ran on RHEL 7.4, was 
equipped with 40 Intel Xeon Gold 6148 2.4GHz CPUs and had
187GB of memory, 94GB of tmpfs and a Dell 120GB M.2 SATA SSD as local disk. 
A 698TB Lustre server was accessible on each node through a 100Gb/s 
Infiniband network. The apparent write bandwidth of the local disk, 
Lustre file system and tmpfs were measured by sequentially writing various numbers
of image blocks containing random intensities, to avoid 
caching effects (measure\_bandwidth.py script in folder 
benchmark\_scripts). They are reported in Table~\ref{table:bdwdths}.

\begin{table}
\centering
\begin{tabular}{c|c}
\rowcolor{headcolor}
Data location & Measured write bandwidths (MB/s)\\
\hline
tmpfs                 & 1377.18 \\
Local disk ($\delta$) & 193.64  \\
Lustre   ($\Delta$)   & 504.03 \\
\end{tabular}
\caption{Measured bandwidths}
\label{table:bdwdths}
\end{table}

% Dedicated Compute nodes with SLURM
% Lustre server

@@

\subsection{Datasets} % 1/4 of a page (half a col)

% BigBrain
% Some BIDS dataset for fmriprep
We used BigBrain~\cite{amunts2013bigbrain}, a 75GB 40$\mu$m isotropic 
histological image of a 65-year-old male's brain. The BigBrain was selected due 
to being the only one of its kind, as there does not yet exist a 
higher-resolution image of a human brain. Moreover, there currently exists a 
lack of standardized tools for processing the BigBrain as a consequence 
of its size. To examine the effects processing BigBrain has on the Big Data 
strategies, we partitioned the full $3845\times3015\times3470$ voxel image into 30 ($5\times3\times2$)
chunks, 125 ($5\times5\times5$) chunks and 750 ($5\times15\times10$) chunks. Additionally, the full 
image was also split in half ($769\times603\times347$ voxels) and the half image was 
partitioned into 125 chunks.

Processing large images is only considered to be part of the Big Data problem 
in neuroscience. The other problem is processing large MRI datasets, that is,
datasets consisting of many small brain images belonging to various different 
subjects. Although we have not explored explicitly the processing of large MRI
datasets, the 75GB BigBrain is within the size ballpark~\cite{van2014human} of 
MRI datasets commonly processed in today's studies.

As it is possible that both small and large data may need to be processed using
the same analysis pipeline, we 
wanted to determine the effects of the data management strategies on small 
data as well. For this, we selected a 12MB T1W image belonging to 
subject 1 of OpenNeuro's 
\href{https://openneuro.org/datasets/ds000001/versions/00006/file-display/sub-01:anat:sub-01_T1w.nii.gz}{ds000001 
dataset version 6}. In order to be able to split it into 125 chunks, it 
was necessary to zero-pad the image to the dimensions
$165\times200\times200$ voxels, which subsequently increased the total image size to 13MB.

\subsection{Applications} % 1 page
\begin{algorithm}\caption{Incrementation}\label{alg:incrementation}
    \begin{algorithmic}[1]
    \Input
    \Desc{$x$}{a sleep delay in seconds}
    \Desc{$n$}{a number of iterations}
    \Desc{$C$}{a set of image chunks}
    \Desc{$fs$}{filesystem to write to (mem, tmpfs, local disk, Lustre).}
    \EndInput
    \ForEach{$chunk \in C$}
    \State read $chunk$ from Lustre
    \For{$i \in [1, n]$}
        \State $chunk\gets chunk+1$
        \State sleep $x$
        \If{$i < n$}
        \State save $chunk$ to $fs$
        \EndIf
    \EndFor
    \State write $chunk$ to Lustre
    \EndFor
\end{algorithmic}
\end{algorithm}

% Incrementation (avoid cache effects in binarization)
In order to investigate how the different strategies impact processing, we 
selected a simple incrementation pipeline (Algorithm~\ref{alg:incrementation}) 
that consisted exclusively of map 
stages. A series of map-only stages would enable us to evaluate the effects of
in-memory computing when data locality is preserved. Incrementation was 
selected 
over other applications, such as binarization, as it ensured that a new image
was created at each step (i.e. no caching effects within the executing 
application). Each partitioned chunk was incremented by 1, in parallel, by
the processing engines. As incrementing the images is a relatively quick 
process, we added a sleep delay to the incrementation stage in order to 
increase task duration as required. The incremented chunks would be either 
maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
Lustre (Spark and Nipype). Should more than a single iteration be requested, 
the increment chunks would be incremented again and saved to the same storage. 
This would repeat until the number of requested iterations had elapsed. In all 
conditions, the first input chunks would be reads from Lustre and the final 
output chunks would be saved to Lustre. We chose to perform our initial 
reads and final writes on Lustre as a multi-tenant clusters' compute node's 
local storage, if 
available, is generally temporary and only available for the duration of the 
allocation.

\subsection{Experiments}

We conducted 4 experiments in which we varied (1) the number of 
iterations in the application ($n$ in 
Algorithm~\ref{alg:incrementation}), (2) the task duration ($x$), (3) 
the chunk size for a constant image size, (4) the total image size. To 
evaluate the page-cache model, experiment conditions fell in different 
regions of Equations~\ref{eq:page-cache-inequality} 
and~\ref{eq:page-cache-sharedfs}, as summarized in 
Table~\ref{table:experiments}. Among the 16 nodes available, 1 was 
dedicated to the Spark master and driver, and the remaining 15 were 
used as compute nodes. Since data-locality is not normally preserved in 
Nipype (a new SLURM allocation is requested for each processed task), 
we instrumented Nipype to ensure data locality. That is, for each chunk 
partition, we requested a SLURM allocation to process the entire 
pipeline in parallel, using Nipype's MultiProc scheduler, on a given 
node. This was possible as no communication was required between the 
processed chunks.
